---
title: "Prediction Project red"
author: "George Saunderson"
date: "5/8/2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---



```{r}
library(dplyr)
library(caret)
library(randomForest)
library(e1071)
library(ggplot2)
library(lattice)
setwd("/Users/geosau_3/Documents/USD/Spring 2018/ECON 386 - Big Data & Business/Project 2")
```
Loads dplyr, caret, randomForest, e1071, ggplot2, and lattice for use with rf analysis. *Assumes all packages are already installed* If NOT installed, use "install.packages("dplyr"),install.packages("caret"), install.packages("randomForest"), install.packages("e1071"), install.packages("ggplot2"), and install.packages("lattice") all on separate lines before loading them.


```{r}
dataset_training <- read.csv("/Users/geosau_3/Documents/USD/Spring 2018/ECON 386 - Big Data & Business/Project 2/training.csv", header = TRUE)
dataset_testing <- read.csv("/Users/geosau_3/Documents/USD/Spring 2018/ECON 386 - Big Data & Business/Project 2/testing.csv", header = TRUE)
```
Simply imports the data with titles for the beginning steps of the data.

```{r}
dataset_training[c(12:17,20,23,26,69:74,87:92,95,98,101,125:130,133,136,139)] <- NULL
dataset_training$X <- NULL
dataset_training_cleanF <- dataset_training[ ,colSums(is.na(dataset_training)) == 0]
dataset_testing[c(12:17,20,23,26,69:74,87:92,95,98,101,125:130,133,136,139)] <- NULL
dataset_testing$X <- NULL
dataset_testing_cleanF <- dataset_testing[ ,colSums(is.na(dataset_testing)) == 0]
```
Removes unnecessary rows and columns from both test and train datasets


```{r}
set.seed(0386)
training_data_index <- createDataPartition(dataset_training_cleanF$classe, p=0.70, list=FALSE)
test_trainingdata_1 <- dataset_training_cleanF[-training_data_index,]
training_data <- dataset_training_cleanF[training_data_index, ]
```
It separates the training data at a 70-30 split, 70% of the training data will be used for training and 30% of the training data with be used for testing. Sets seet at 5 (group number = 5).

```{r}
set.seed(0386)
test_trainingdata_2 <- sample(1:nrow(test_trainingdata_1), size = 0.5*nrow(test_trainingdata_1))
tune_data <- test_trainingdata_1[test_trainingdata_2, ]
test_data <- test_trainingdata_1[-test_trainingdata_2, ]
```
Separates the 30% of training data that will be used for testing into two parts, creating a 70-15-15 split where 70 = training, 15 = tuning, and 15 = testing. Sets seet at 5 (group number = 5).

```{r}
dim(dataset_testing)
dim(dataset_testing_cleanF)
dim(dataset_training)
dim(dataset_training_cleanF)
dim(training_data)
dim(test_trainingdata_1)
length(as.numeric(test_trainingdata_2))
dim(tune_data)
dim(test_data)
```
Shows the dimensions of each set of data so far.

```{r}
set.seed(0386)
control <- trainControl(method = "cv", 20)
metric <- "Accuracy"
```
Splits dataset into 20 different parts for crossvalidation. Sets seet at 0386.

```{r}
set.seed(0386)
rf_classe <- train(classe ~ ., data = training_data, na.action = na.omit, method = "rf", metric = metric, trControl = control)
CART_classe <- train(classe ~ ., data = training_data, na.action = na.omit, method = "rpart", metric = metric, trControl = control)
results_classe <- resamples(list(cart=CART_classe, rf=rf_classe))
summary(results_classe)
```
Creates a random forest model and sets seet at 5 (group number = 5). Displays summary to make decision on optimal model.

```{r}
prediction_rf <- predict(rf_classe, test_data)
cM_rf <- confusionMatrix(prediction_rf, test_data$classe)
```
Produces both a prediction and the confusion matrix for the given prediction of the test data.

```{r}
summary(rf_classe)
print(rf_classe)
```
Shows the results of the model in graphics.

```{r}
prediction_testFinal <- predict(rf_classe, dataset_testing_cleanF)
```
Prediction on final test dataset.
