---
title: "Prediction Project red"
author: "George Saunderson"
date: "5/8/2018"
output:
  html_document:
    df_print: paged
---



```{r}
library(dplyr)
library(caret)
library(randomForest)
library(e1071)
library(ggplot2)
library(lattice)
setwd("/Users/geosau_3/Documents/USD/Spring 2018/ECON 386 - Big Data & Business/Project 2")
```
Loads dplyr, caret, randomForest, e1071, ggplot2, and lattice for use with rf analysis.


```{r}
dataset_training <- read.csv("/Users/geosau_3/Documents/USD/Spring 2018/ECON 386 - Big Data & Business/Project 2/training.csv")
dataset_testing <- read.csv("/Users/geosau_3/Documents/USD/Spring 2018/ECON 386 - Big Data & Business/Project 2/testing.csv")
```
Simply imports the data with titles for the beginning steps of the data.

```{r}
dataset_training[c(12:17,20,23,26,69:74,87:92,95,98,101,125:130,133,136,139)] <- NULL
dataset_training$X <- NULL
dataset_testing$X <- NULL
```
Removes columns with empty cells only. Does NOT remove columns with NA in the cells. Nullifies irrelavant and unnecessary columns.


```{r}
training_data_index <- createDataPartition(dataset_training$classe, p=0.75, list=FALSE)
test_trainingdata <- dataset_training[-training_data_index,]
training_data <- dataset_training[training_data_index, ]
```
It separates the training data at a 75-25 split, 75% of the training data will be used for training and 25% of the training data with be used for testing.

```{r}
dim(dataset_testing)
dim(dataset_training)
dim(training_data)
dim(test_trainingdata)
```
Shows the dimensions of each set of data so far.

```{r}
control <- trainControl(method = "cv", 20)
metric <- "Accuracy"
```
Splits dataset into 20 different parts for crossvalidation.

```{r}
set.seed(5)
rf_classe <- train(classe ~ ., data = training_data, na.action = na.exclude, method = "rf", metric = metric, trControl = control)

set.seed(5)
svm_classe <- train(classe ~ ., data = training_data, na.action = na.exclude, method = "svmRadial", metric = metric, trControl = control)

set.seed(5)
CART_classe <- train(classe ~ ., data = training_data, na.action = na.exclude, method = "rpart", metric = metric, trControl = control)

set.seed(5)
knn_classe <- train(classe ~ ., data = training_data, na.action = na.exclude, method = "knn", metric = metric, trControl = control)
```
Creates four different models (rf, svm, CART, and knn), two being non-linear and another two being advanced algorithms.

```{r}
results <- resamples(list(rf=rf_classe, svm=svm_classe, cart=CART_classe, knn=knn_classe))
summary(results)
dotplot(results)
```
Shows the results for each model to make a choice for the optimal model by both showing a summary and a dotplot for comparisons. (seed = group # [5])

After seeing the results, rf looks to be the optimal model to proceed with.
```{r}
print(rf_classe)
```
Gives a summary of the rf_classe.

```{r}
predictions <- predict(rf_classe, test_trainingdata)
head(predictions)
```
Shows the first five predictions based on the rf model applied to the test_trainingdata.
